Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-28 Manuscript/manuscript.Rmd
manuscript edits

Multiple random forest classification algorithms [??? What you mean by multiple? Random forest already has multiplicity in it.] were built on domain data extracted from Hi-C data for 2 cell lines $C = \{\textit{GM12878}, \textit{K562}\}$, at four different resolutions $R = \{\textit{10 kb}, \textit{25 kb}, \textit{50 kb}, \textit{100 kb\}}$. [??? Bad spacing, for the whole paragraph. Why you are not using plain and simple Markdown?] For each classifier, we compared model performances using 3 separate predictor types $P = \{\textit{OC}, \textit{OP}, \textit{Distance}\}$ and 4 class balancing techniques $B = \{\textit{None}, \textit{RUS}, \textit{ROS}, \textit{SMOTE}\}$. A system produced from the ensemble framework was defined as: $S = \{\textit{c}, \textit{r}, \textit{p}, \textit{b}\}$, where $c \in C$, $r \in R$, $p \in P$, and $b \in B$. Letting $|X|$ denote the cardinality of each set, we evaluated $|C| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $96 (= 2 \times 4 \times 3 \times 4)$ separate models.

@@

To assess the full [???] severity of class imbalance present when constructing the models, we investigated the relationship between the total number of bins (sample observations) included at each resolution [??? That's mundane, supplemental at best] and the subsequent imbalance ratio (IR). [??? Did you ever defined it?] As expected, the total number of genomic bins decreased as resolution decreased (Figure 4A). Specifically, at lower resolutions, the linear genome was binned into larger units, thereby creating fewer bins. however, the IRs remained relatively stable and consistent across resolutions and cell lines (Figure 4B). We found that there was an average IR of 0.154 (SD=0.022) and 0.137 (SD=0.0181) across resolutions for the GM12878 and K562 cell lines respectively. Here, the low imbalance ratios indicated a wide disparity between the number of genomic bins that overlap with a flanked TAD boundary and the much larger number that did not.

## HIGH CORRRELATIONS IMPLY COLOCALIZATION OF MANY ANNOTATIONS AROUND TAD BOUNDARIES

@@

The high correlations among many of the genomic annotations gave precedent for the use of an elastic-net regularization algorithm in order to suitably reduce the feature space while retaining possibly correlated annotations that could be predictive of TAD boundary formation. For each model created from the ensemble framework, the alpha and lambda were tuned over a grid of values using 10-fold cross validation (Supplementary Table 7).

>>>>

We compared the performances of random forest classification models implemented on TAD boundaries called using the arrowhead algorithm [@durand2016juicebox] from Hi-C data for 2 cell lines $C = \{\textit{GM12878}, \textit{K562}\}$, at four different resolutions $R = \{\textit{10 kb}, \textit{25 kb}, \textit{50 kb}, \textit{100 kb\}}$. [??? Bad spacing, for the whole paragraph. Why you are not using plain and simple Markdown?] For each classifier, we considered 3 separate predictor types as the feature space including $P = \{\textit{OC}, \textit{OP}, \textit{Distance}\}$. To assess the severity of class imbalance 4 re-sampling techniques were implemented prior to modelling, including $B = \{\textit{None}, \textit{RUS}, \textit{ROS}, \textit{SMOTE}\}$. A system produced from the ensemble framework was defined as: $S = \{\textit{c}, \textit{r}, \textit{p}, \textit{b}\}$, where $c \in C$, $r \in R$, $p \in P$, and $b \in B$. Letting $|X|$ denote the cardinality of each set, we evaluated $|C| \times |R| \times |P| \times |B|$ different ensemble systems, totaling $96 (= 2 \times 4 \times 3 \times 4)$ separate models.

@@

Results showed that the imbalance ratios were well below acceptable thresholds for both cell lines, and across the 4 resolutions (Supplementary Table 9 ???SDs will come with more chromosomes). The low imbalance ratios indicated a wide disparity between the number of genomic bins that overlap with a flanked TAD boundary and the much larger number that did not. While resolution certainly played a role in determining the number of observations included in datasets used in downstream modeling, as it determines the width of genomic bins and larger widths (lower resolutions) created less bins, we found that there was very little variation in imbalance ratios across resolution, with ICC=X and Y for GM12878 and K562 respectively (Supplementary Figure 16 ???ICC values will come with more chromosomes).

<!--
## HIGH CORRRELATIONS IMPLY COLOCALIZATION OF MANY ANNOTATIONS AROUND TAD BOUNDARIES

@@

The high correlations among many of the genomic annotations gave precedent for the use of an elastic-net regularization algorithm in order to suitably reduce the feature space while retaining possibly correlated annotations that could be predictive of TAD boundary formation. For each model created from the ensemble framework, the alpha and lambda were tuned over a grid of values using 10-fold cross validation (Supplementary Table 7).
-->
